---
title: "get_grooves_robust_loess"
author: "Kiegan Rice"
date: "5/7/2018"
output: pdf_document
---

## Hamby44

**Note: This is the same intro as the "get_grooves_quadratic_explo.R" file - we are making the same adjustments.**  


We will first take a look at the Hamby 44 Rescan data. There are lists of groove predictions (using Eric's smoothing method "rollapply") and groove manual identifications (done by Dr. Hofmann).

`cccata`, within the Hamby 44 Rescan data, contains x-y-value formatted data. 


```{r packages-and-data, echo = F, message = F, warning = F}
library(devtools)
#install.packages("tidyverse")
#library(tidyverse)
library(tidyr)
library(dplyr)
library(purrr)
library(ggplot2)

#devtools::install_github("CSAFE-ISU/bulletr")
#library(bulletr)


#hamby44 <- readRDS("/home/ricek/hamby44.rda")
hamby44 <- readRDS("//opt/hamby44/hamby44.rda")
#hamby44
hamby44[1:6,]
```

There are a few changes we want to make to the `hamby44` data before we fit things.  

The first is that there are a few lands that are `NA` values, so we just need to filter those out before we can do anything else. It looks like it only happens in two cases, so we are still left with 208 viable lands.    

```{r rm-empty-lands, message = F, warning = F}
hamby44 <- hamby44 %>% filter(!is.na(crosscuts))
```


The second is that since `hamby44` has 10 crosscuts for each land, we will collapse those into one by averaging them. Then hamby44 will have an extra column, ccdata_avg, which is a data frame for each land consisting of the columns `y` and `value`, but here `value` is the averaged values. We will also create a `value_std` which is the averaged value standardized down to 0 if the lowest value is not 0. This can be seen below; the 10 crosscuts are plotted as points, in black. The averaged values are plotted as a red line, and the standardized average values are plotted as a green line. While this is just done for the first land in the data set, and there is not much of a difference here, for some it will be a more pronounced difference.  




```{r avg-and-std-values, echo = F, message = F, warning = F}

hamby44 <- hamby44 %>% mutate(ccdata_avg = purrr::map(ccdata, .f = function(dframe){
  dframe <- dframe %>% group_by(y) %>% summarise(value = mean(value, na.rm = T))
  dframe <- as.data.frame(dframe)
  check_min <- min(dframe$value[!is.na(dframe$value)])
  dframe <- dframe %>% mutate(value_std = value - check_min)
  return(dframe)
}))

#hamby44

ggplot() + geom_point(data = hamby44$ccdata[[1]], aes(x = y, y = value)) + 
  geom_line(data = hamby44$ccdata_avg[[1]], aes(x = y, y = value), color = "red") + 
  geom_line(data = hamby44$ccdata_avg[[1]], aes(x = y, y = value_std), colour = "green") + 
  theme_bw()
```


**Here is where things start to change...**  

We are interested in looking at the residuals from a robust LOESS fit (this changes from the robust linear model fit we previously were considering). There are many instances of a bad linear fit due to a truly non-quadratic bullet land shape with robust linear models, so we want to consider robust LOESS, which appears to do a much better job of both addressing outlier values in the edges and capturing the shape of the bullet land. Thus, we will fit this model and calculate the residuals that occur in order to later use them in assessment. So here, we will calculate the residuals from the robust linear model (`absresid`, `resid`) AND the robust LOESS model (`rlo_absresid`, `rlo_resid`).  

Note that we will do all of this with the `value_std` column from `ccdata_avg`, so we will need to make sure we use `value_std` when we actually run the function later. It calls `value` as it is defined. *Maybe add a parameter to the function that lets you define what the value column is called.*  


```{r calculate-residuals-rlm-rloess, echo = F, warning = F, message = F}
library(locfit)
hamby44 <- hamby44 %>% mutate(ccdata_w_resid = purrr::map(ccdata_avg, .f = function(bullet){
  lm0 <- MASS::rlm(value_std~poly(y,2), data=bullet, maxit=100)
  bullet$pred <- predict(lm0, newdata=bullet)

  bullet$absresid <- with(bullet, abs(value_std-pred))
  bullet$resid <- with(bullet, value_std-pred)
  
  
  robust_loess_fit <- locfit.robust(value_std~y, data = bullet, alpha = 1, kern = "tcub")
  bullet$rlo_pred <- predict(robust_loess_fit, newdata = bullet)
  
  bullet$rlo_absresid <- with(bullet, abs(value_std-rlo_pred))
  bullet$rlo_resid <- with(bullet, value_std-rlo_pred)
  return(bullet)
}), rlm_r2 = purrr::map_dbl(ccdata_w_resid, .f = function(bullet){
  r2 <- (cor(bullet$value_std, bullet$pred, use = "complete.obs"))^2
  return(r2)
}), rlo_r2 = purrr::map_dbl(ccdata_w_resid, .f = function(bullet){
  r2 <- (cor(bullet$value_std, bullet$rlo_pred, use = "complete.obs"))^2
  return(r2)
})
)

hamby44 %>% ggplot() + geom_point(aes(x = rlm_r2, y = rlo_r2)) + theme_bw() + geom_abline(intercept = 0, slope = 1, colour = "red")


#bullet_plot <- function(i){
#  bullet <- hamby44$ccdata_w_resid[[i]]
#  bullet %>% ggplot() + geom_point(aes(x = y, y = value_std)) + theme_bw() + geom_line(aes(x = y, y = rlo_pred), colour ="green")
#}
```


Now that we have all the values we are interested in comparing to calculated, we can define several different functions we had previously considered using for groove identification.  


### get_grooves_rollapply  

here we aren't going to actually define/run the function, there are just "grooves_pred" in the hamby44 datafile.  

```{r calculate-sum_resid_rolls, echo = F, warning = F, message = F}
sum_resid_right_roll <- function(datafile, which_bullet){
  data <- datafile$ccdata_w_resid[[which_bullet]]
  man_groove <- datafile$grooves[[which_bullet]]$groove[2]
  pred_groove <- datafile$grooves_pred[[which_bullet]]$groove[2]
  data <- data %>% filter(between(y, min(man_groove, pred_groove), max(man_groove, pred_groove)))
  return(sum(data$resid, na.rm = T))
}

sum_rresid_right_roll <- function(datafile, which_bullet){
  data <- datafile$ccdata_w_resid[[which_bullet]]
  man_groove <- datafile$grooves[[which_bullet]]$groove[2]
  pred_groove <- datafile$grooves_pred[[which_bullet]]$groove[2]
  data <- data %>% filter(between(y, min(man_groove, pred_groove), max(man_groove, pred_groove)))
  return(sum(data$rlo_resid, na.rm = T))
}

hamby44$rdiff_resid_roll <- 0
hamby44$rdiff_rresid_roll <- 0
for (i in 1:nrow(hamby44)){
  hamby44$rdiff_resid_roll[i] <- sum_resid_right_roll(datafile = hamby44, which_bullet = i)
  hamby44$rdiff_rresid_roll[i] <- sum_rresid_right_roll(datafile = hamby44, which_bullet = i)
}


sum_resid_left_roll <- function(datafile, which_bullet){
  data <- datafile$ccdata_w_resid[[which_bullet]]
  man_groove <- datafile$grooves[[which_bullet]]$groove[1]
  pred_groove <- datafile$grooves_pred[[which_bullet]]$groove[1]
  data <- data %>% filter(between(y, min(man_groove, pred_groove), max(man_groove, pred_groove)))
  return(sum(data$resid, na.rm = T))
}
sum_rresid_left_roll <- function(datafile, which_bullet){
  data <- datafile$ccdata_w_resid[[which_bullet]]
  man_groove <- datafile$grooves[[which_bullet]]$groove[1]
  pred_groove <- datafile$grooves_pred[[which_bullet]]$groove[1]
  data <- data %>% filter(between(y, min(man_groove, pred_groove), max(man_groove, pred_groove)))
  return(sum(data$rlo_resid, na.rm = T))
}

hamby44$ldiff_resid_sum_roll <- 0
hamby44$ldiff_rresid_sum_roll <- 0
for (i in 1:nrow(hamby44)){
  hamby44$ldiff_resid_sum_roll[i] <- sum_resid_left_roll(datafile = hamby44, which_bullet = i)
  hamby44$ldiff_rresid_sum_roll[i] <- sum_rresid_left_roll(datafile = hamby44, which_bullet = i)
}

```

### get_grooves_quadratic  

```{r define-get_grooves_quadratic, message = F, warning = F}
get_grooves_quadratic <- function(bullet, adjust) {
  lm0 <- MASS::rlm(value_std~poly(y,2), data=bullet, maxit=100)
  bullet$pred <- predict(lm0, newdata=bullet)

  bullet$absresid <- with(bullet, abs(value_std-pred))
  bullet$absresid90 <- with(
    bullet, absresid>4*median(bullet$absresid, na.rm=TRUE))

  groove <- range(filter(bullet, !absresid90)$y) + c(adjust, -adjust)
  
  plot <- bullet %>% ggplot(aes(x = y, y = value_std)) + geom_line(size = .5) + theme_bw() +
    geom_vline(xintercept=groove[1], colour = "blue") +
    geom_vline(xintercept=groove[2], colour = "blue") 
  
  return(list(groove = groove, plot = plot))
}

```

```{r run-get_grooves_quadratic, message = F, warning = F}

hamby44 <- hamby44 %>% mutate(grooves_pred_quadratic = purrr::map(ccdata_avg, .f = function(x){
  get_grooves_quadratic(bullet = x, adjust = 0)$groove
}))

grooves_pred_quadratic <- hamby44$grooves_pred_quadratic
hamby44_eval <- readRDS("../data/hamby44/hamby44_eval.rda")
hamby44_eval$grooves_pred_quadratic <- grooves_pred_quadratic
saveRDS(hamby44_eval, "../data/hamby44/hamby44_eval.rda")
```


```{r calculate-sum_resid_quads, message = F, warning = F, echo = F}

sum_resid_right <- function(datafile, which_bullet){
  data <- datafile$ccdata_w_resid[[which_bullet]]
  man_groove <- datafile$grooves[[which_bullet]]$groove[2]
  pred_groove <- datafile$grooves_quad_pred[[which_bullet]][2]
  data <- data %>% filter(between(y, min(man_groove, pred_groove), max(man_groove, pred_groove)))
  return(sum(data$resid, na.rm = T))
}

sum_rresid_right <- function(datafile, which_bullet){
  data <- datafile$ccdata_w_resid[[which_bullet]]
  man_groove <- datafile$grooves[[which_bullet]]$groove[2]
  pred_groove <- datafile$grooves_quad_pred[[which_bullet]][2]
  data <- data %>% filter(between(y, min(man_groove, pred_groove), max(man_groove, pred_groove)))
  return(sum(data$rlo_resid, na.rm = T))
}

hamby44$rdiff_resid_sum <- 0
hamby44$rdiff_rresid_sum <- 0
for (i in 1:nrow(hamby44)){
  hamby44$rdiff_resid_sum[i] <- sum_resid_right(datafile = hamby44, which_bullet = i)
  hamby44$rdiff_rresid_sum[i] <- sum_rresid_right(datafile = hamby44, which_bullet = i)
}


sum_resid_left <- function(datafile, which_bullet){
  data <- datafile$ccdata_w_resid[[which_bullet]]
  man_groove <- datafile$grooves[[which_bullet]]$groove[1]
  pred_groove <- datafile$grooves_quad_pred[[which_bullet]][1]
  data <- data %>% filter(between(y, min(man_groove, pred_groove), max(man_groove, pred_groove)))
  return(sum(data$resid, na.rm = T))
}
sum_rresid_left <- function(datafile, which_bullet){
  data <- datafile$ccdata_w_resid[[which_bullet]]
  man_groove <- datafile$grooves[[which_bullet]]$groove[1]
  pred_groove <- datafile$grooves_quad_pred[[which_bullet]][1]
  data <- data %>% filter(between(y, min(man_groove, pred_groove), max(man_groove, pred_groove)))
  return(sum(data$rlo_resid, na.rm = T))
}

hamby44$ldiff_resid_sum <- 0
hamby44$ldiff_rresid_sum <- 0
for (i in 1:nrow(hamby44)){
  hamby44$ldiff_resid_sum[i] <- sum_resid_left(datafile = hamby44, which_bullet = i)
  hamby44$ldiff_rresid_sum[i] <- sum_rresid_left(datafile = hamby44, which_bullet = i)
}
```

### get_grooves_quadratic_pos_resid  

```{r define-get_grooves_quadratic_pos_resid, message = F, warning = F}
get_grooves_quadratic_pos_resid <- function(bullet, adjust) {
  lm0 <- MASS::rlm(value_std~poly(y,2), data=bullet, maxit=100)
  bullet$pred <- predict(lm0, newdata=bullet)

  #bullet$absresid <- with(bullet, abs(value-pred))
  bullet$resid <- with(bullet, value_std-pred)
  #bullet$absresid_div <- with(bullet, absresid/value)
  bullet$resid_div <- with(bullet, resid/value_std)
  #bullet %>% ggplot() + geom_point(aes(x = y, y = resid)) + theme_bw()
  #bullet %>% ggplot() + geom_point(aes(x = y, y = resid_div)) + theme_bw()
  
  bullet$resid_cutoff <- with (bullet, resid_div > 0)
  groove <- range(filter(bullet, !resid_cutoff)$y) + c(adjust, -adjust)
  
  plot <- bullet %>% ggplot(aes(x = y, y = value_std)) + geom_line(size = .5) + theme_bw() +
    geom_vline(xintercept=groove[1], colour = "blue") +
    geom_vline(xintercept=groove[2], colour = "blue") 
  
  return(list(groove = groove, plot = plot))
}
```


```{r run-get_grooves_quadratic_pos_resid, echo = F, warning = F, message = F}
hamby44 <- hamby44 %>% mutate(grooves_quad_pred_pos = purrr::map(ccdata_avg, .f = function(x){
  get_grooves_quadratic_pos_resid(bullet = x, adjust = 0)$groove
}))
```

```{r calculate-sum_resid_quads_pos, echo = F, warning = F, message = F}

sum_resid_right_pos <- function(datafile, which_bullet){
  data <- datafile$ccdata_w_resid[[which_bullet]]
  man_groove <- datafile$grooves[[which_bullet]]$groove[2]
  pred_groove <- datafile$grooves_quad_pred_pos[[which_bullet]][2]
  data <- data %>% filter(between(y, min(man_groove, pred_groove), max(man_groove, pred_groove)))
  return(sum(data$resid, na.rm = T))
}
sum_rresid_right_pos <- function(datafile, which_bullet){
  data <- datafile$ccdata_w_resid[[which_bullet]]
  man_groove <- datafile$grooves[[which_bullet]]$groove[2]
  pred_groove <- datafile$grooves_quad_pred_pos[[which_bullet]][2]
  data <- data %>% filter(between(y, min(man_groove, pred_groove), max(man_groove, pred_groove)))
  return(sum(data$rlo_resid, na.rm = T))
}

hamby44$rdiff_resid_sum_pos <- 0
hamby44$rdiff_rresid_sum_pos <- 0
for (i in 1:nrow(hamby44)){
  hamby44$rdiff_resid_sum_pos[i] <- sum_resid_right_pos(datafile = hamby44, which_bullet = i)
  hamby44$rdiff_rresid_sum_pos[i] <- sum_rresid_right_pos(datafile = hamby44, which_bullet = i)
}


sum_resid_left_pos <- function(datafile, which_bullet){
  data <- datafile$ccdata_w_resid[[which_bullet]]
  man_groove <- datafile$grooves[[which_bullet]]$groove[1]
  pred_groove <- datafile$grooves_quad_pred_pos[[which_bullet]][1]
  data <- data %>% filter(between(y, min(man_groove, pred_groove), max(man_groove, pred_groove)))
  return(sum(data$resid, na.rm = T))
}
sum_rresid_left_pos <- function(datafile, which_bullet){
  data <- datafile$ccdata_w_resid[[which_bullet]]
  man_groove <- datafile$grooves[[which_bullet]]$groove[1]
  pred_groove <- datafile$grooves_quad_pred_pos[[which_bullet]][1]
  data <- data %>% filter(between(y, min(man_groove, pred_groove), max(man_groove, pred_groove)))
  return(sum(data$rlo_resid, na.rm = T))
}

hamby44$ldiff_resid_sum_pos <- 0
hamby44$ldiff_rresid_sum_pos <- 0
for (i in 1:nrow(hamby44)){
  hamby44$ldiff_resid_sum_pos[i] <- sum_resid_left_pos(datafile = hamby44, which_bullet = i)
  hamby44$ldiff_rresid_sum_pos[i] <- sum_resid_left_pos(datafile = hamby44, which_bullet = i)
}

```


### get_grooves_mad  

```{r define_get-grooves-mad, echo = F, message = F, warning = F}
get_grooves_mad <- function(bullet, adjust, multiplier){
  lm0 <- MASS::rlm(value_std~poly(y,2), data=bullet, maxit=100)
  bullet$pred <- predict(lm0, newdata=bullet)

  bullet$absresid <- with(bullet, abs(value_std-pred))
  #bullet$resid <- with(bullet, value_std-pred)
  bullet$absresid_div <- with(bullet, absresid/value_std)
  #bullet$resid_div <- with(bullet, resid/value_std)
  #bullet %>% ggplot() + geom_point(aes(x = y, y = resid)) + theme_bw()
  #bullet %>% ggplot() + geom_point(aes(x = y, y = resid_div)) + theme_bw()
  mad1 <- mad(bullet$absresid, na.rm = T)
  bullet$resid_cutoff <- with(bullet, absresid > mad1*multiplier)
  groove <- range(filter(bullet, !resid_cutoff)$y) + c(adjust, -adjust)
  
  plot <- bullet %>% ggplot(aes(x = y, y = value_std)) + geom_line(size = .5) + theme_bw() +
    geom_vline(xintercept=groove[1], colour = "blue") +
    geom_vline(xintercept=groove[2], colour = "blue") 
  
  return(list(groove = groove, plot = plot))
}

```

```{r calculate_get-grooves-mad1, echo = F, message = F, warning = F}
hamby44 <- hamby44 %>% mutate(grooves_mad1 = purrr::map(ccdata_avg, .f = function(x){
  get_grooves_mad(bullet = x, adjust = 0, multiplier = 1)$groove
}))
```

```{r calculate-sum_resid_mads, echo = F, warning = F, message = F}
sum_resid_right_m1 <- function(datafile, which_bullet){
  data <- datafile$ccdata_w_resid[[which_bullet]]
  man_groove <- datafile$grooves[[which_bullet]]$groove[2]
  pred_groove <- datafile$grooves_mad1[[which_bullet]][2]
  data <- data %>% filter(between(y, min(man_groove, pred_groove), max(man_groove, pred_groove)))
  return(sum(data$resid, na.rm = T))
}

sum_rresid_right_m1 <- function(datafile, which_bullet){
  data <- datafile$ccdata_w_resid[[which_bullet]]
  man_groove <- datafile$grooves[[which_bullet]]$groove[2]
  pred_groove <- datafile$grooves_mad1[[which_bullet]][2]
  data <- data %>% filter(between(y, min(man_groove, pred_groove), max(man_groove, pred_groove)))
  return(sum(data$rlo_resid, na.rm = T))
}


hamby44$rdiff_resid_sum_m1 <- 0
hamby44$rdiff_rresid_sum_m1 <- 0
for (i in 1:nrow(hamby44)){
  hamby44$rdiff_resid_sum_m1[i] <- sum_resid_right_m1(datafile = hamby44, which_bullet = i)
  hamby44$rdiff_rresid_sum_m1[i] <- sum_rresid_right_m1(datafile = hamby44, which_bullet = i)
}


sum_resid_left_m1 <- function(datafile, which_bullet){
  data <- datafile$ccdata_w_resid[[which_bullet]]
  man_groove <- datafile$grooves[[which_bullet]]$groove[1]
  pred_groove <- datafile$grooves_mad1[[which_bullet]][1]
  data <- data %>% filter(between(y, min(man_groove, pred_groove), max(man_groove, pred_groove)))
  return(sum(data$resid, na.rm = T))
}

sum_rresid_left_m1 <- function(datafile, which_bullet){
  data <- datafile$ccdata_w_resid[[which_bullet]]
  man_groove <- datafile$grooves[[which_bullet]]$groove[1]
  pred_groove <- datafile$grooves_mad1[[which_bullet]][1]
  data <- data %>% filter(between(y, min(man_groove, pred_groove), max(man_groove, pred_groove)))
  return(sum(data$rlo_resid, na.rm = T))
}

hamby44$ldiff_resid_sum_m1 <- 0
hamby44$ldiff_rresid_sum_m1 <- 0
for (i in 1:nrow(hamby44)){
  hamby44$ldiff_resid_sum_m1[i] <- sum_resid_left_m1(datafile = hamby44, which_bullet = i)
  hamby44$ldiff_rresid_sum_m1[i] <- sum_rresid_left_m1(datafile = hamby44, which_bullet = i)
}
```


### get_grooves_mad2

```{r calculate_get-grooves-mad2, echo = F, message = F, warning = F}
hamby44 <- hamby44 %>% mutate(grooves_mad2 = purrr::map(ccdata_avg, .f = function(x){
  get_grooves_mad(bullet = x, adjust = 0, multiplier = 2)$groove
}))
```  


### predicting groove sides  

```{r explore-groove-locations, echo = F, warning = F, message = F}

hamby44$left_groove <- 0
hamby44$right_groove <- 0
for(i in 1:nrow(hamby44)){
  hamby44$left_groove[i] <- hamby44$grooves[[i]]$groove[1]
  hamby44$right_groove[i] <- hamby44$grooves[[i]]$groove[2]
}

summary(hamby44$left_groove) # Left grooves, from 15.48 to 543.74
summary(hamby44$right_groove) # Right grooves, from 1872 to 2351

hamby44 %>% ggplot() + geom_histogram(aes(x = left_groove), fill = "salmon")  + 
  geom_histogram(aes(x = right_groove), fill = "turquoise") + 
  labs(x = "Groove Locations") + theme_bw()
```

Now, let's take a look at summing up residuals from 0 to the highest left groove location, and from the lowest right groove location up to the max y value.  

```{r calculate-left-right-sums-means, echo = F, warning = F, message = F}
hamby44 <- hamby44 %>% mutate(left_resid_sum = purrr::map_dbl(ccdata_w_resid, .f = function(bullet){
  bullet <- bullet %>% filter(y < 600) 
  sum(bullet$resid, na.rm = T)
}), 
right_resid_sum = purrr::map_dbl(ccdata_w_resid, .f = function(bullet){
  bullet <- bullet %>% filter(y > 1800)
  sum(bullet$resid, na.rm = T)
}),
left_rlo_resid_sum = purrr::map_dbl(ccdata_w_resid, .f = function(bullet){
  bullet <- bullet %>% filter(y < 600)
  sum(bullet$rlo_resid, na.rm = T)
}), 
right_rlo_resid_sum = purrr::map_dbl(ccdata_w_resid, .f = function(bullet){
  bullet <- bullet %>% filter(y > 1800)
  sum(bullet$rlo_resid, na.rm = T)
}))

hamby44 <- hamby44 %>% mutate(left_resid_mean = purrr::map_dbl(ccdata_w_resid, .f = function(bullet){
  bullet <- bullet %>% filter(y < 600) 
  mean(bullet$resid, na.rm = T)
}), 
right_resid_mean = purrr::map_dbl(ccdata_w_resid, .f = function(bullet){
  bullet <- bullet %>% filter(y > 1800)
  mean(bullet$resid, na.rm = T)
}),
left_rlo_resid_mean = purrr::map_dbl(ccdata_w_resid, .f = function(bullet){
  bullet <- bullet %>% filter(y < 600)
  mean(bullet$rlo_resid, na.rm = T)
}), 
right_rlo_resid_mean = purrr::map_dbl(ccdata_w_resid, .f = function(bullet){
  bullet <- bullet %>% filter(y > 1800)
  mean(bullet$rlo_resid, na.rm = T)
}))

hamby44 %>% ggplot() + geom_histogram(aes(x = left_resid_sum), fill = "salmon", alpha = 0.3) + geom_histogram(aes(x = right_resid_sum), fill = "turquoise", alpha = 0.3) + theme_bw() 

hamby44 %>% ggplot() + geom_histogram(aes(x = left_rlo_resid_sum), fill = "salmon", alpha = 0.3) + geom_histogram(aes(x = right_rlo_resid_sum), fill = "turquoise", alpha = 0.3) + theme_bw()

hamby44 %>% ggplot() + geom_point(aes(x = left_resid_sum, y = left_rlo_resid_sum)) + theme_bw() + geom_abline(intercept = 0, slope = 1, color = "red")
hamby44 %>% ggplot() + geom_point(aes(x = right_resid_sum, y = right_rlo_resid_sum)) + theme_bw() + geom_abline(intercept = 0, slope = 1, color = "red")

hamby44 %>% ggplot() + geom_point(aes(x = left_resid_mean, y = left_rlo_resid_mean)) + theme_bw() + geom_abline(intercept = 0, slope = 1, color = "red")
hamby44 %>% ggplot() + geom_point(aes(x = right_resid_mean, y = right_rlo_resid_mean)) + theme_bw() + geom_abline(intercept = 0, slope = 1, color = "red")

## Based on the plot, let's go with a cutoff of residual higher than 1000 means there is a groove there for the right side... but I have no idea for the right side!! Let's see if we can get some sort of ground truth...  



```









### define whether there is or is not a groove  

```{r define-is-groove, echo = F, warning = F, message = F}

hamby44$is_left_groove <- c(1, 1, 1, 1, 1, #5
                            1, 1, 1, 1, 0, #10
                            1, 1, 1, 1, 1, #15
                            1, 1, 1, 1, 1, #20
                            1, 1, 1, 1, 1, #25
                            1, 1, 1, 1, 1, #30
                            1, 0, 1, 1, 1, #35
                            1, 1, 1, 1, 1, #40
                            1, 1, 1, 1, 1, #45
                            1, 1, 1, 1, 1, #50
                            1, 1, 1, 1, 1, #55
                            1, 1, 0, 1, 1, #60
                            1, 1, 1, 1, 1, #65
                            1, 1, 1, 1, 1, #70
                            1, 1, 1, 1, 1, #75
                            1, 1, 1, 1, 1, #80
                            1, 1, 1, 1, 1, #85
                            1, 1, 1, 1, 1, #90
                            1, 1, 1, 1, 1, #95
                            1, 1, 1, 0, 1, #100
                            1, 1, 1, 1, 1, #105
                            1, 1, 1, 1, 1, #110
                            1, 1, 1, 1, 1, #115
                            1, 1, 1, 1, 1, #120
                            1, 1, 1, 1, 1, #125
                            1, 1, 1, 1, 1, #130
                            1, 1, 1, 1, 1, #135
                            1, 1, 1, 1, 1, #140
                            1, 1, 1, 1, 1, #145
                            1, 1, 1, 1, 1, #150
                            1, 1, 1, 1, 1, #155
                            1, 1, 1, 1, 1, #160
                            1, 1, 1, 1, 1, #165
                            1, 1, 1, 1, 1, #170
                            1, 1, 1, 1, 1, #175
                            1, 1, 1, 1, 1, #180
                            1, 1, 1, 1, 1, #185
                            1, 1, 1, 1, 1, #190
                            1, 1, 1, 1, 1, #195
                            1, 1, 1, 1, 1, #200
                            1, 1, 1, 1, 1, #205
                            1, 1, 1)

hamby44$is_right_groove <- c(1, 1, 0, 0, 0, #5 
                             1, 0, 0, 0, 1, #10
                             0, 0, 1, 1, 0, #15
                             0, 1, 1, 1, 0, #20
                             0, 0, 0, 1, 0, #25
                             1, 0, 0, 0, 0, #30
                             1, 1, 1, 0, 1, #35
                             0, 1, 1, 1, 0, #40
                             0, 0, 0, 1, 0, #45
                             1, 0, 1, 0, 0, #50
                             1, 0, 0, 1, 0, #55
                             0, 0, 1, 1, 0, #60
                             0, 1, 0, 0, 1, #65
                             1, 0, 1, 1, 1, #70
                             0, 0, 1, 0, 1, #75
                             0, 1, 0, 1, 0, #80
                             0, 1, 0, 1, 1, #85
                             1, 0, 1, 0, 1, #90
                             0, 0, 0, 1, 1, #95
                             1, 0, 1, 1, 1, #100
                             0, 0, 1, 1, 0, #105
                             0, 0, 1, 0, 0, #110
                             0, 1, 1, 1, 0, #115
                             0, 1, 0, 1, 0, #120
                             1, 1, 1, 0, 0, #125
                             0, 1, 1, 1, 1, #130
                             1, 1, 0, 1, 1, #135
                             1, 1, 0, 0, 0, #140
                             0, 0, 0, 0, 0, #145
                             0, 1, 1, 0, 0, #150
                             1, 1, 0, 1, 0, #155
                             0, 1, 0, 0, 0, #160
                             0, 0, 0, 0, 1, #165
                             1, 0, 1, 0, 0, #170
                             1, 1, 0, 0, 0, #175
                             0, 0, 1, 0, 1, #180
                             0, 1, 0, 1, 0, #185
                             1, 1, 1, 1, 0, #190
                             1, 1, 1, 0, 1, #195
                             1, 1, 1, 0, 1, #200
                             0, 1, 0, 1, 0, #205
                             1, 0, 1)


```



```{r, echo = F, warning = F, message = F}

hamby44 %>% filter(!is.na(is_left_groove)) %>% ggplot() + geom_histogram(aes(x = left_rlo_resid_sum, fill = factor(is_left_groove))) + theme_bw()

hamby44 %>% filter(!is.na(is_right_groove)) %>% ggplot() + geom_histogram(aes(x = right_rlo_resid_sum, fill = factor(is_right_groove)), binwidth = 200) + theme_bw()


hamby44 %>% ggplot() + geom_point(aes(x = left_rlo_resid_sum, y = left_groove, colour = factor(is_left_groove))) + theme_bw()

hamby44 %>% ggplot() + geom_point(aes(x = right_rlo_resid_sum, y = right_groove, colour = factor(is_right_groove))) + theme_bw()


hamby44 %>% filter(!is.na(is_left_groove)) %>% ggplot() + geom_histogram(aes(x = left_rlo_resid_mean, fill = factor(is_left_groove))) + theme_bw()

hamby44 %>% filter(!is.na(is_right_groove)) %>% ggplot() + geom_histogram(aes(x = right_rlo_resid_mean, fill = factor(is_right_groove))) + theme_bw()


hamby44 %>% ggplot() + geom_point(aes(x = left_rlo_resid_mean, y = left_groove, colour = factor(is_left_groove))) + theme_bw()

hamby44 %>% ggplot() + geom_point(aes(x = right_rlo_resid_mean, y = right_groove, colour = factor(is_right_groove))) + theme_bw()
```



### get_grooves_robust_loess  



```{r explore-lowess, echo = F, warning = F, message = F}


bullet <- hamby44$ccdata_w_resid[[132]]
bullet_filter <- subset(bullet, !is.na(value_std))
loess_fit <- fitted(loess(value_std~y, data = bullet_filter, span = 0.9))

## let's try locfit.robust from the locfit package instead
#install.packages("locfit")
library(locfit)

#wts <- ifelse(bullet_filter$y < 100 | bullet_filter$y > 1800, 0, 1)
robust_loess_fit <- locfit.robust(value_std~y, data = bullet_filter, kern = "tcub", alpha = 1)
r_loess_fit <- predict(robust_loess_fit, newdata = bullet_filter)
## this looks like a better "robust" fit, and looks fairly promising at that!  


ggplot() + geom_point(data = bullet_filter, aes(x = y, y = value_std)) + 
  #geom_line(aes(x = bullet_filter$y, y = loess_fit), colour = "red") + 
  geom_line(aes(x = bullet_filter$y, y = r_loess_fit), colour = "green") + 
  theme_bw()



```



```{r define-get_grooves_robust_loess_compare, echo = F, warning = F, message = F}
get_grooves_robust_loess_compare <- function(bullet, adjust){
  bullet$loess_fit <- predict(loess(value_std~y, data = bullet, span = 1), newdata = bullet)
  bullet$robust_loess_fit <- predict(locfit.robust(value_std~y, data = bullet, alpha = 1), newdata = bullet)
  
  left <- bullet %>% filter(y < 600) ## this sums up residuals on the left 1/4 or so of the land
  sumleft <- sum(left$loess_fit - left$robust_loess_fit, na.rm = T)
  right <- bullet %>% filter(y > 1800) ## this sums up residuals on the right 1/4 or so of the land  
  sumright <- sum(right$loess_fit - right$robust_loess_fit, na.rm = T)
  
  
  ## can't find a working "Trim" function for numerics, so I'm going to use a little workaround I found  
  
  ## let's first trim down the left side - we almost always need to trim it down.  
  groove_l <- 0
  while(sumleft > 300){
    bullet <- bullet[(round(0.01 * length(bullet$y))) : (length(bullet$y)),]
    loess_new <- loess(value_std~y, data = bullet, span = 1)
    rloess_new <- locfit.robust(value_std~y, data = bullet, alpha = 1)
    #lm_new <- MASS::rlm(value_std~poly(y,2), data=bullet, maxit=100)
    bullet$pred_new <- predict(loess_new, newdata = bullet)
    bullet$rpred_new <- predict(rloess_new, newdata = bullet)
    bullet$diff_new <- with(bullet, pred_new-rpred_new)
    sumleft <- sum(bullet[(bullet$y < 600),]$diff_new, na.rm = T)
    groove_l <- min(bullet$y)
    #redefine sumleft here
  }
  
  groove_r <- max(bullet$y)
  while(sumright > 300){
    bullet <- bullet[1:(round(0.99*length(bullet$y))),]
    loess_new <- loess(value_std~y, data = bullet, span = 1)
    rloess_new <- locfit.robust(value_std~y, data = bullet, alpha = 1)
    bullet$pred_new <- predict(loess_new, newdata = bullet)
    bullet$rpred_new <- predict(rloess_new, newdata = bullet)
    #lm_new <- MASS::rlm(value_std~poly(y,2), data = bullet, maxit=100)
    #bullet$pred_new <- predict(lm_new, newdata = bullet)
    bullet$diff_new <- with(bullet, pred_new-rpred_new)
    sumright <- sum(bullet[(bullet$y > 1800),]$diff_new, na.rm = T)
    groove_r <- max(bullet$y)
  }
  
  
  groove <- c(groove_l, groove_r)
  
  
  #plot <- bullet %>% ggplot() + geom_line(aes(x = y, y = value_std), size = .5) + theme_bw() +
  #  geom_vline(xintercept=groove[1], colour = "blue") +
  #  geom_vline(xintercept=groove[2], colour = "blue") 
  return(list(groove = groove))
  #return(list(groove = groove, plot = plot))
}


## Look at AVERAGE residual - this will stop getting pulled up as much once the shoulders are gone! 
## this also takes into account how many points are there (gets rid of NA issues)  

```

```{r calculate-get_grooves_robust_loess_compare, echo = F, warning = F, message = F}
hamby44 <- hamby44 %>% mutate(grooves_rl1 = purrr::map(ccdata_avg, .f = function(x){
  get_grooves_robust_loess_compare(bullet = x, adjust = 0)$groove
}))
```


```{r calculate_sum_resids_rl1, echo = F, warning = F, message = F}

sum_resid_right_rl1 <- function(datafile, which_bullet){
  data <- datafile$ccdata_w_resid[[which_bullet]]
  man_groove <- datafile$grooves[[which_bullet]]$groove[2]
  pred_groove <- datafile$grooves_rl1[[which_bullet]][2]
  data <- data %>% filter(between(y, min(man_groove, pred_groove), max(man_groove, pred_groove)))
  return(sum(data$resid, na.rm = T))
}

## name you want the column to be called in the dataframe goes here <- function() 
sum_rresid_right_rl1 <- function(datafile, which_bullet){
  data <- datafile$ccdata_w_resid[[which_bullet]]
  man_groove <- datafile$grooves[[which_bullet]]$groove[2]
  pred_groove <- datafile$grooves_rl1[[which_bullet]][2]
  data <- data %>% filter(between(y, min(man_groove, pred_groove), max(man_groove, pred_groove)))
  return(sum(data$rlo_resid, na.rm = T))
}


hamby44$rdiff_resid_sum_rl1 <- 0
hamby44$rdiff_rresid_sum_rl1 <- 0
for (i in 1:nrow(hamby44)){
  hamby44$rdiff_resid_sum_rl1[i] <- sum_resid_right_rl1(datafile = hamby44, which_bullet = i)
  hamby44$rdiff_rresid_sum_rl1[i] <- sum_rresid_right_rl1(datafile = hamby44, which_bullet = i)
}


sum_resid_left_rl1 <- function(datafile, which_bullet){
  data <- datafile$ccdata_w_resid[[which_bullet]]
  man_groove <- datafile$grooves[[which_bullet]]$groove[1]
  pred_groove <- datafile$grooves_rl1[[which_bullet]][1]
  data <- data %>% filter(between(y, min(man_groove, pred_groove), max(man_groove, pred_groove)))
  return(sum(data$resid, na.rm = T))
}

sum_rresid_left_rl1 <- function(datafile, which_bullet){
  data <- datafile$ccdata_w_resid[[which_bullet]]
  man_groove <- datafile$grooves[[which_bullet]]$groove[1]
  pred_groove <- datafile$grooves_rl1[[which_bullet]][1]
  data <- data %>% filter(between(y, min(man_groove, pred_groove), max(man_groove, pred_groove)))
  return(sum(data$rlo_resid, na.rm = T))
}

hamby44$ldiff_resid_sum_rl1 <- 0
hamby44$ldiff_rresid_sum_rl1 <- 0
for (i in 1:nrow(hamby44)){
  hamby44$ldiff_resid_sum_rl1[i] <- sum_resid_left_rl1(datafile = hamby44, which_bullet = i)
  hamby44$ldiff_rresid_sum_rl1[i] <- sum_rresid_left_rl1(datafile = hamby44, which_bullet = i)
}


```


Now let's try two variants on the robust loess - one where we look at the mean residuals and one where we look at a straight cutoff value.  


```{r define-get_grooves_robust_loess_mean, echo = F, warning = F, message = F}
get_grooves_robust_loess_mean <- function(bullet, adjust){
  #bullet$loess_fit <- predict(loess(value_std~y, data = bullet, span = 1), newdata = bullet)
  bullet$robust_loess_fit <- predict(locfit.robust(value_std~y, data = bullet, alpha = 1), newdata = bullet)
  
  bullet$resid1 <- bullet$value_std - bullet$robust_loess_fit
  left <- bullet %>% filter(y < 600) ## this sums up residuals on the left 1/4 or so of the land
  meanleft <- mean(left$resid1, na.rm = T)
  right <- bullet %>% filter(y > 1800) ## this sums up residuals on the right 1/4 or so of the land  
  meanright <- mean(right$resid1, na.rm = T)
  
  
  ## can't find a working "Trim" function for numerics, so I'm going to use a little workaround I found  
  
  ## let's first trim down the left side - we almost always need to trim it down.  
  groove_l <- 0
  num_points <- sum(!is.na(left$value_std))
  while(meanleft > 1 & num_points > 10){
    bullet <- bullet[(round(0.01 * length(bullet$y))) : (length(bullet$y)),]
    #loess_new <- loess(value_std~y, data = bullet, span = 1)
    rloess_new <- locfit.robust(value_std~y, data = bullet, alpha = 1)
    #lm_new <- MASS::rlm(value_std~poly(y,2), data=bullet, maxit=100)
    #bullet$pred_new <- predict(loess_new, newdata = bullet)
    bullet$rpred_new <- predict(rloess_new, newdata = bullet)
    bullet$resid_new <- with(bullet, value_std-rpred_new)
    meanleft <- mean(bullet[(bullet$y < 600),]$resid_new, na.rm = T)
    num_points <- sum(!is.na(bullet[(bullet$y < 600),]$value_std))
    groove_l <- min(bullet$y)
    #redefine sumleft here
  }
  
  groove_r <- max(bullet$y)
  num_points <- sum(!is.na(right$value_std))
  while(meanright > 1 & num_points > 10){
    bullet <- bullet[1:(round(0.99*length(bullet$y))),]
    #loess_new <- loess(value_std~y, data = bullet, span = 1)
    rloess_new <- locfit.robust(value_std~y, data = bullet, alpha = 1)
    #bullet$pred_new <- predict(loess_new, newdata = bullet)
    bullet$rpred_new <- predict(rloess_new, newdata = bullet)
    #lm_new <- MASS::rlm(value_std~poly(y,2), data = bullet, maxit=100)
    #bullet$pred_new <- predict(lm_new, newdata = bullet)
    bullet$resid_new <- with(bullet, value_std-rpred_new)
    meanright <- mean(bullet[(bullet$y > 1800),]$resid_new, na.rm = T)
    num_points <- sum(!is.na(bullet[(bullet$y > 1800),]$value_std))
    groove_r <- max(bullet$y)
  }
  
  
  groove <- c(groove_l, groove_r)
  
  
  #plot <- bullet %>% ggplot() + geom_line(aes(x = y, y = value_std), size = .5) + theme_bw() +
  #  geom_vline(xintercept=groove[1], colour = "blue") +
  #  geom_vline(xintercept=groove[2], colour = "blue") 
  return(list(groove = groove))
  #return(list(groove = groove, plot = plot))
}


## Look at AVERAGE residual - this will stop getting pulled up as much once the shoulders are gone! 
## this also takes into account how many points are there (gets rid of NA issues)  

```

```{r calculate-get_grooves_robust_loess_mean, echo = F, warning = F, message = F}
hamby44 <- hamby44 %>% mutate(grooves_rl_mean = purrr::map(ccdata_avg, .f = function(x){
  get_grooves_robust_loess_mean(bullet = x, adjust = 0)$groove
}))
```


```{r calculate_sum_resids_rl_mean, echo = F, warning = F, message = F}

sum_resid_right_rlmean <- function(datafile, which_bullet){
  data <- datafile$ccdata_w_resid[[which_bullet]]
  man_groove <- datafile$grooves[[which_bullet]]$groove[2]
  pred_groove <- datafile$grooves_rl_mean[[which_bullet]][2]
  data <- data %>% filter(between(y, min(man_groove, pred_groove), max(man_groove, pred_groove)))
  return(sum(data$resid, na.rm = T))
}

sum_rresid_right_rlmean <- function(datafile, which_bullet){
  data <- datafile$ccdata_w_resid[[which_bullet]]
  man_groove <- datafile$grooves[[which_bullet]]$groove[2]
  pred_groove <- datafile$grooves_rl_mean[[which_bullet]][2]
  data <- data %>% filter(between(y, min(man_groove, pred_groove), max(man_groove, pred_groove)))
  return(sum(data$rlo_resid, na.rm = T))
}

hamby44$rdiff_resid_sum_rlmean <- 0
hamby44$rdiff_rresid_sum_rlmean <- 0
for (i in 1:nrow(hamby44)){
  hamby44$rdiff_resid_sum_rlmean[i] <- sum_resid_right_rlmean(datafile = hamby44, which_bullet = i)
  hamby44$rdiff_rresid_sum_rlmean[i] <- sum_rresid_right_rlmean(datafile = hamby44, which_bullet = i)
}


sum_resid_left_rlmean <- function(datafile, which_bullet){
  data <- datafile$ccdata_w_resid[[which_bullet]]
  man_groove <- datafile$grooves[[which_bullet]]$groove[1]
  pred_groove <- datafile$grooves_rl_mean[[which_bullet]][1]
  data <- data %>% filter(between(y, min(man_groove, pred_groove), max(man_groove, pred_groove)))
  return(sum(data$resid, na.rm = T))
}

sum_rresid_left_rlmean <- function(datafile, which_bullet){
  data <- datafile$ccdata_w_resid[[which_bullet]]
  man_groove <- datafile$grooves[[which_bullet]]$groove[1]
  pred_groove <- datafile$grooves_rl_mean[[which_bullet]][1]
  data <- data %>% filter(between(y, min(man_groove, pred_groove), max(man_groove, pred_groove)))
  return(sum(data$rlo_resid, na.rm = T))
}

hamby44$ldiff_resid_sum_rlmean <- 0
hamby44$ldiff_rresid_sum_rlmean <- 0
for (i in 1:nrow(hamby44)){
  hamby44$ldiff_resid_sum_rlmean[i] <- sum_resid_left_rlmean(datafile = hamby44, which_bullet = i)
  hamby44$ldiff_rresid_sum_rlmean[i] <- sum_rresid_left_rlmean(datafile = hamby44, which_bullet = i)
}


```



Try raw residual cutoff values with robust linear model! 

```{r}
#plot_resids <- function(i){
#  bullet <- hamby44$ccdata_w_resid[[i]]
#  bullet %>% ggplot() + geom_point(aes(x = y, y = abs(rlo_resid))) + theme_bw() + geom_hline(yintercept = 0, color = "red") + #geom_hline(yintercept = 4*median(abs(rlo_resid), na.rm = T), color = "blue")
#}
```

```{r define-get_grooves_robust_loess_cutoff, echo = F, warning = F, message = F}
get_grooves_robust_loess_cutoff <- function(bullet, adjust) {
  robust_loess_fit <- locfit.robust(value_std~y, data = bullet, alpha = 1)
  bullet$pred <- predict(robust_loess_fit, newdata = bullet)
  #bullet$pred <- predict(lm0, newdata=bullet)

  bullet$absresid <- with(bullet, abs(value_std-pred))
  bullet$resid <- with(bullet, value_std-pred)
  #bullet$absresid_div <- with(bullet, absresid/value)
  bullet$resid_div <- with(bullet, resid/value_std)
  #bullet %>% ggplot() + geom_point(aes(x = y, y = resid)) + theme_bw()
  #bullet %>% ggplot() + geom_point(aes(x = y, y = resid_div)) + theme_bw()
  cutoff <- 2*median(bullet$absresid, na.rm = T)
  bullet$resid_cutoff <- with(bullet, absresid > cutoff)
  groove <- range(filter(bullet, !resid_cutoff)$y) + c(adjust, -adjust)
  
  plot <- bullet %>% ggplot(aes(x = y, y = value_std)) + geom_line(size = .5) + theme_bw() +
    geom_vline(xintercept=groove[1], colour = "blue") +
    geom_vline(xintercept=groove[2], colour = "blue") 
  
  return(list(groove = groove, plot = plot))
}
```


```{r calculate-get_grooves_robust_loess_cutoff, echo = F, warning = F, message = F}
hamby44 <- hamby44 %>% mutate(grooves_pred_rlo = purrr::map(ccdata_avg, .f = function(x){
  get_grooves_robust_loess_cutoff(bullet = x, adjust = 0)$groove
}))


grooves_pred_rlo <- hamby44$grooves_pred_rlo
hamby44_eval <- readRDS("../data/hamby44/hamby44_eval.rda")
hamby44_eval$grooves_pred_rlo <- grooves_pred_rlo
saveRDS(hamby44_eval, "../data/hamby44/hamby44_eval.rda")
```



```{r calculate_sum_resids_rlcut, echo = F, warning = F, message = F}

sum_resid_right_rlcut <- function(datafile, which_bullet){
  data <- datafile$ccdata_w_resid[[which_bullet]]
  man_groove <- datafile$grooves[[which_bullet]]$groove[2]
  pred_groove <- datafile$grooves_rl_cutoff[[which_bullet]][2]
  data <- data %>% filter(between(y, min(man_groove, pred_groove), max(man_groove, pred_groove)))
  return(sum(data$resid, na.rm = T))
}

sum_rresid_right_rlcut <- function(datafile, which_bullet){
  data <- datafile$ccdata_w_resid[[which_bullet]]
  man_groove <- datafile$grooves[[which_bullet]]$groove[2]
  pred_groove <- datafile$grooves_rl_cutoff[[which_bullet]][2]
  data <- data %>% filter(between(y, min(man_groove, pred_groove), max(man_groove, pred_groove)))
  return(sum(data$rlo_resid, na.rm = T))
}

hamby44$rdiff_resid_sum_rlcut <- 0
hamby44$rdiff_rresid_sum_rlcut <- 0
for (i in 1:nrow(hamby44)){
  hamby44$rdiff_resid_sum_rlcut[i] <- sum_resid_right_rlcut(datafile = hamby44, which_bullet = i)
  hamby44$rdiff_rresid_sum_rlcut[i] <- sum_rresid_right_rlcut(datafile = hamby44, which_bullet = i)
}


sum_resid_left_rlcut <- function(datafile, which_bullet){
  data <- datafile$ccdata_w_resid[[which_bullet]]
  man_groove <- datafile$grooves[[which_bullet]]$groove[1]
  pred_groove <- datafile$grooves_rl_cutoff[[which_bullet]][1]
  data <- data %>% filter(between(y, min(man_groove, pred_groove), max(man_groove, pred_groove)))
  return(sum(data$resid, na.rm = T))
}

sum_rresid_left_rlcut <- function(datafile, which_bullet){
  data <- datafile$ccdata_w_resid[[which_bullet]]
  man_groove <- datafile$grooves[[which_bullet]]$groove[1]
  pred_groove <- datafile$grooves_rl_cutoff[[which_bullet]][1]
  data <- data %>% filter(between(y, min(man_groove, pred_groove), max(man_groove, pred_groove)))
  return(sum(data$rlo_resid, na.rm = T))
}

hamby44$ldiff_resid_sum_rlcut <- 0
hamby44$ldiff_rresid_sum_rlcut <- 0
for (i in 1:nrow(hamby44)){
  hamby44$ldiff_resid_sum_rlcut[i] <- sum_resid_left_rlcut(datafile = hamby44, which_bullet = i)
  hamby44$ldiff_rresid_sum_rlcut[i] <- sum_rresid_left_rlcut(datafile = hamby44, which_bullet = i)
}


```



```{r plots_for_afte, echo = F, warning = F, message = F}
cbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#999999")

plot_resids_grooves <- function(i){
  bullet <- hamby44$ccdata_w_resid[[i]]
  left_groove <- hamby44$grooves_rl_cutoff[[i]][1]
  right_groove <- hamby44$grooves_rl_cutoff[[i]][2]
  bullet %>% ggplot() + geom_point(aes(x = y, y = rlo_resid)) + 
  geom_hline(yintercept = 0, color = cbPalette[2]) + 
  geom_vline(xintercept = left_groove, color = "red") + 
  geom_vline(xintercept = right_groove, color = "red") + 
  theme_bw() + labs(x = "Y Value", y = "Residual Value")
}
plot_resids_grooves(68)



highlight_plot <- function(i){
  bullet <- hamby44$ccdata_w_resid[[i]]
  ggplot() + geom_point(data = bullet, aes(x = y, y = rlo_resid)) + 
  geom_vline(xintercept = hamby44$grooves_rl_cutoff[[i]][1], colour = "blue") + 
  geom_vline(xintercept = hamby44$grooves_rl_cutoff[[i]][2], colour = "blue") + 
  geom_hline(yintercept = 0, colour = "red") +
  #geom_vline(xintercept = hamby44$grooves[[20]]$groove[1], colour = "green")  + 
  #geom_vline(xintercept = hamby44$grooves[[20]]$groove[2], colour = "green") + 
  labs(x = "Y Values", y = "Height Values", title = "Predicting Groove Presence - Residuals") +
  geom_rect(aes(ymin = min(bullet$resid, na.rm = T), 
                ymax = max(bullet$resid, na.rm = T),
                xmax = max(hamby44$grooves[[i]]$groove[1], hamby44$grooves_rl_cutoff[[i]][1]), 
                xmin = min(hamby44$grooves[[i]]$groove[1], hamby44$grooves_rl_cutoff[[i]][1])), fill = "yellow", alpha = 0.3) +
  geom_rect(aes(ymin = min(bullet$resid, na.rm = T), 
                ymax = max(bullet$resid, na.rm = T),
                xmax = max(hamby44$grooves[[i]]$groove[2], hamby44$grooves_rl_cutoff[[i]][2]), 
                xmin = min(hamby44$grooves[[i]]$groove[2], hamby44$grooves_rl_cutoff[[i]][2])), fill = "yellow", alpha = 0.3) + 
  #geom_line(data = hamby44$ccdata_w_resid[[20]], aes(x = y, y = pred), colour = "red") + 
  theme_bw()
}
highlight_plot(68)
```




```{r}
resids_dframe_2 <- data.frame(roll_right_q = abs(hamby44$rdiff_resid_roll), roll_left_q = abs(hamby44$ldiff_resid_sum_roll), 
                            roll_right_rlo = abs(hamby44$rdiff_rresid_roll), roll_left_rlo = abs(hamby44$ldiff_rresid_sum_roll), 
                             quad_right_q = abs(hamby44$rdiff_resid_sum), quad_left_q = abs(hamby44$ldiff_resid_sum), 
                             quad_right_rlo = abs(hamby44$rdiff_rresid_sum), quad_left_rlo = abs(hamby44$ldiff_rresid_sum),
                             rlcut_right_q = abs(hamby44$rdiff_resid_sum_rlcut), rlcut_left_q = abs(hamby44$ldiff_resid_sum_rlcut),
                             rlcut_right_rlo = abs(hamby44$rdiff_rresid_sum_rlcut), rlcut_left_rlo = abs(hamby44$ldiff_rresid_sum_rlcut)
                             )

apply(resids_dframe_2, MARGIN = 2, mean)
apply(resids_dframe_2, MARGIN = 2, median)

head(resids_dframe_2)

#resids_dframe$method <- factor(resids_dframe$method, levels = rev(levels(resids_dframe$method)))
resids_dframe_2 <- resids_dframe_2 %>% gather(method, residuals_sum)

#resids_dframe$method <- factor(resids_dframe$method,
#    levels = c('roll_left', 'quad_left', 'pos_left', 'mad1_left', 'mad2_left', 'roll_right', 'quad_right', 'pos_right', 'mad1_right', 'mad2_right'),ordered = TRUE)

#resids_dframe %>% ggplot() + geom_boxplot(aes(x = method, y = residuals_sum)) + 
#  theme_bw() + labs(x = "Method and Groove", y = "Residuals") + scale_x_discrete(labels = c("Rollapply, Left", "RLM, Left", "Rollapply, Right", "RLM, Right"))

#resids_dframe  %>% select(-quad_right, -quad_left) %>% gather(method, residuals_sum) %>% filter(residuals_sum < 2000) %>% 
#  ggplot() + geom_boxplot(aes(x = factor(method), y = residuals_sum)) + theme_bw()

resids_dframe_2 <- resids_dframe_2 %>% mutate(method = as.character(method), Method = unlist(purrr::map(method, .f = function(x){
  strsplit(x, "_")[[1]][1]
})), GrooveSide = unlist(purrr::map(method, .f = function(x){
  strsplit(x, "_")[[1]][2]
})), 
ResidType = unlist(purrr::map(method, .f = function(x){
  strsplit(x, "_")[[1]][3]
})))


resids_dframe_2 %>% filter(ResidType == "rlo") %>% mutate(Method = factor(Method, levels = levels(as.factor(Method))[c(3,1,2)], 
                                         labels = c("Rollapply", "Robust Linear Model", 
                                                    "Robust Loess"))) %>% 
  ggplot() + geom_boxplot(aes(x = GrooveSide, y = residuals_sum, fill = Method)) + 
  theme_bw() + labs(y = "Sum of Residuals") + scale_fill_manual(values = cbPalette)


resids_dframe_2 %>% filter(ResidType == "rlo", residuals_sum < 10000) %>% mutate(Method = factor(Method, levels = levels(as.factor(Method))[c(3,1,2)], 
                                         labels = c("Rollapply", "Robust Linear Model", 
                                                    "Robust Loess"))) %>% 
  ggplot() + geom_boxplot(aes(x = GrooveSide, y = residuals_sum, fill = Method)) + 
  theme_bw() + labs(y = "Sum of Residuals") + scale_fill_manual(values = cbPalette)
```


```{r compare-get-grooves-methods, echo = F, warning = F, message = F}

resids_dframe <- data.frame(roll_right_q = abs(hamby44$rdiff_resid_roll), roll_left_q = abs(hamby44$ldiff_resid_sum_roll), 
                            roll_right_rlo = abs(hamby44$rdiff_rresid_roll), roll_left_rlo = abs(hamby44$ldiff_rresid_sum_roll), 
                             quad_right_q = abs(hamby44$rdiff_resid_sum), quad_left_q = abs(hamby44$ldiff_resid_sum), 
                             quad_right_rlo = abs(hamby44$rdiff_rresid_sum), quad_left_rlo = abs(hamby44$ldiff_rresid_sum),
                             pos_right_q = abs(hamby44$rdiff_resid_sum_pos), pos_left_q = abs(hamby44$ldiff_resid_sum_pos),
                             pos_right_rlo = abs(hamby44$rdiff_rresid_sum_pos), pos_left_rlo = abs(hamby44$ldiff_rresid_sum_pos),
                             mad1_right_q = abs(hamby44$rdiff_resid_sum_m1), mad1_left_q = abs(hamby44$ldiff_resid_sum_m1),
                             mad1_right_rlo = abs(hamby44$rdiff_rresid_sum_m1), mad1_left_rlo = abs(hamby44$ldiff_rresid_sum_m1),
                             rl1_right_q = abs(hamby44$rdiff_resid_sum_rl1), rl1_left_q = abs(hamby44$ldiff_resid_sum_rl1),
                             rl1_right_rlo = abs(hamby44$rdiff_rresid_sum_rl1), rl1_left_rlo = abs(hamby44$ldiff_rresid_sum_rl1),
                             rlmean_right_q = abs(hamby44$rdiff_resid_sum_rlmean), rlmean_left_q = abs(hamby44$ldiff_resid_sum_rlmean),
                             rlmean_right_rlo = abs(hamby44$rdiff_rresid_sum_rlmean), rlmean_left_rlo = abs(hamby44$ldiff_rresid_sum_rlmean),
                             rlcut_right_q = abs(hamby44$rdiff_resid_sum_rlcut), rlcut_left_q = abs(hamby44$ldiff_resid_sum_rlcut),
                             rlcut_right_rlo = abs(hamby44$rdiff_rresid_sum_rlcut), rlcut_left_rlo = abs(hamby44$ldiff_rresid_sum_rlcut)
                             )

apply(resids_dframe, MARGIN = 2, mean)
apply(resids_dframe, MARGIN = 2, median)

head(resids_dframe)

#resids_dframe$method <- factor(resids_dframe$method, levels = rev(levels(resids_dframe$method)))
resids_dframe <- resids_dframe %>% gather(method, residuals_sum)

#resids_dframe$method <- factor(resids_dframe$method,
#    levels = c('roll_left', 'quad_left', 'pos_left', 'mad1_left', 'mad2_left', 'roll_right', 'quad_right', 'pos_right', 'mad1_right', 'mad2_right'),ordered = TRUE)

#resids_dframe %>% ggplot() + geom_boxplot(aes(x = method, y = residuals_sum)) + 
#  theme_bw() + labs(x = "Method and Groove", y = "Residuals") + scale_x_discrete(labels = c("Rollapply, Left", "RLM, Left", "Rollapply, Right", "RLM, Right"))

#resids_dframe  %>% select(-quad_right, -quad_left) %>% gather(method, residuals_sum) %>% filter(residuals_sum < 2000) %>% 
#  ggplot() + geom_boxplot(aes(x = factor(method), y = residuals_sum)) + theme_bw()

resids_dframe <- resids_dframe %>% mutate(method = as.character(method), Method = unlist(purrr::map(method, .f = function(x){
  strsplit(x, "_")[[1]][1]
})), GrooveSide = unlist(purrr::map(method, .f = function(x){
  strsplit(x, "_")[[1]][2]
})), 
ResidType = unlist(purrr::map(method, .f = function(x){
  strsplit(x, "_")[[1]][3]
})))


resids_dframe %>% mutate(Method = factor(Method, levels = levels(as.factor(Method))[c(7,3,2,1,4,6,5)], 
                                         labels = c("Rollapply", "Quadratic", "Robust Linear Model", 
                                                    "Median Absolute Deviation", "Robust Loess", "Robust Loess Mean Resids", "Robust Loess Cutoff"))) %>% 
  ggplot() + geom_boxplot(aes(x = GrooveSide, y = residuals_sum, fill = Method)) + 
  theme_bw() + labs(y = "Sum of Residuals") + facet_wrap(~ResidType)


resids_dframe %>% 
  mutate(Method = factor(Method, levels = levels(as.factor(Method))[c(7,3,2,1,4,6,5)], 
                         labels = c("Rollapply", "Quadratic", "Robust Linear Model", "Median Absolute Deviation", "Robust Loess", "Robust Loess Mean Resids", "Robust Loess Cutoff")))  %>%
  filter(residuals_sum < 1000) %>% 
  ggplot() + geom_boxplot(aes(x = GrooveSide, y = residuals_sum, fill = Method)) + 
  theme_bw() + labs(y = "Sum of Residuals") + facet_wrap(~ResidType)



```








```{r}

hamby44 %>% ggplot() + geom_point(aes(x = y, y  ))



#hamby44$ccdata_w_resid[[11]] %>% ggplot() + geom_point(aes(x = y, y = value_std)) +
#  geom_line(aes(x = y, y = pred), colour = "red") + 
#  geom_line(aes(x = y, y = rlo_pred), colour = "green") + 
#  theme_bw()


plot_b4 <- hamby44$ccdata_w_resid[[4]] %>% ggplot() + geom_point(aes(x = y, y = value_std)) +
  geom_line(aes(x = y, y = pred), colour = "red") + 
  geom_line(aes(x = y, y = rlo_pred), colour = "green") + 
  theme_bw()

plot_b5 <- hamby44$ccdata_w_resid[[5]] %>% ggplot() + geom_point(aes(x = y, y = value_std)) +
  geom_line(aes(x = y, y = pred), colour = "red") + 
  geom_line(aes(x = y, y = rlo_pred), colour = "green") + 
  theme_bw()

plot_b8 <- hamby44$ccdata_w_resid[[8]] %>% ggplot() + geom_point(aes(x = y, y = value_std)) +
  geom_line(aes(x = y, y = pred), colour = "red") + 
  geom_line(aes(x = y, y = rlo_pred), colour = "green") + 
  theme_bw()

plot_b11 <- hamby44$ccdata_w_resid[[11]] %>% ggplot() + geom_point(aes(x = y, y = value_std)) +
  geom_line(aes(x = y, y = pred), colour = "red") + 
  geom_line(aes(x = y, y = rlo_pred), colour = "green") + 
  theme_bw()


plot_b6 <- hamby44$ccdata_w_resid[[6]] %>% ggplot() + geom_point(aes(x = y, y = value_std)) +
  geom_line(aes(x = y, y = pred), colour = "red") + 
  geom_line(aes(x = y, y = rlo_pred), colour = "green") + 
  theme_bw()
plot_b10 <- hamby44$ccdata_w_resid[[10]] %>% ggplot() + geom_point(aes(x = y, y = value_std)) +
  geom_line(aes(x = y, y = pred), colour = "red") + 
  geom_line(aes(x = y, y = rlo_pred), colour = "green") + 
  theme_bw()


library(gridExtra)

grid.arrange(plot_b4, plot_b5, plot_b8, plot_b11, ncol = 2)

grid.arrange(plot_b6, plot_b10, ncol =2 )

```
### Note: will need to re-define "sum_resid_right", "sum_resid_left"... these are how we are summing up the residuals between manual ID's and predicted grooves. This is a really gross way to do it, also... there must be a better way!  




