---
title: "Evaluating and Saving Predictions"
output: html_document
---

This document is intended to streamline the process of collaborating and evaluating groove predictions on the groove identification project.  

I have created a datafile, `hamby44_eval`, to assist us in assessing our predicted groove locations. I have removed the `x3p`, `grooves`, `processed`, and `proc_smooth` columns from hamby44 to reduce the size of the file we are actually working with. This leaves us with the following: 

1. Metadata about the lands: `source`, `barrel`, `bullet`, `land`, `PARTITION_ID`, `crosscuts`.  
2. The land data averaged across crosscuts (`ccdata_w_resid`). This is most likely what you want to start with for modeling.  
    - Values shifted down so min value occurs at 0 (`value_std`)
    - Predictions and residuals from robust LOESS fit (`rlo_pred`, `rlo_resid`, `rlo_absresid`)
3. Manually identified "ground truth" locations for the grooves: `left_groove`, `right_groove`.  
4. Rollapply method groove predictions: `grooves_pred`.  
5. "Scores" for rollapply method predictions: `score_left_rollapply`, `score_right_rollapply`.  


The idea here is that we can add new prediction columns as we make them.  

The first important step is in your document that you are using to run code/make predictions, your output should be a list that contains predictions for each land. So the list should have 208 elements (in the Hamby44 set), and each item in the list should have [1] and [2], the predictions for the left and right groove. If you follow the `purrr::map` structure to each land and return a list: `return(list(groove = groove, plot = plot))` (if you don't do the plot part, just do `return(list(groove = groove))`. Just make sure whatever function you are applying to the lands returns a list. Then you want to grab the $groove part of the list when you actually apply the function to a bunch of lands. 

You can then save your predictions as an added column on the `hamby44_eval` file:  

```{r example-save}
#grooves_pred_"method" <- hamby44$grooves_pred_"method"
#hamby44_eval <- readRDS("../data/hamby44/hamby44_eval.rda") ## the "../" at the beginning takes you from a file in the "development" folder of the grooves repository
#hamby44_eval$grooves_pred_"method" <- grooves_pred_"method"
#saveRDS(hamby44_eval, "../data/hamby44/hamby44_eval.rda")
```

If you have trouble getting your results into the right format, let me know and I can help you out!  


THEN, we can read it in and run through the following code to get results and save these "scores" for the predictions to the hamby44_eval datafile.  

```{r read-hamby44_eval}
hamby44_eval <- readRDS("data/hamby44/hamby44_eval.rda")
```

```{r define-prediction_scores}
prediction_scores <- function(dataset, method){
  colname <- paste0("grooves_pred_", method)
  new_colname <- paste0("score_left_", method)
  new_colname2 <- paste0("score_right_", method)
  dataset[,new_colname] <- 0
  dataset[,new_colname2] <- 0
  for(i in 1:nrow(dataset)){
    bullet <- dataset[i,]
    left_pred <- bullet[,colname][[1]][[1]][1]
    left_groove <- as.numeric(bullet[,"left_groove"])
    dat <- bullet$ccdata_w_resid[[1]]
    dat <- dat %>% filter(between(y, min(left_pred, left_groove), max(left_pred, left_groove)))
    dataset[i,new_colname] <- sum(dat$rlo_resid, na.rm = T)
    
    right_pred <- bullet[,colname][[1]][[1]][2]
    right_groove <- as.numeric(bullet[,"right_groove"])
    dat2 <- bullet$ccdata_w_resid[[1]]
    dat2 <- dat2 %>% filter(between(y, min(right_pred, right_groove), max(right_pred, right_groove)))
    dataset[i, new_colname2] <- sum(dat2$rlo_resid, na.rm = T)
  }
  return(dataset)
}
```

```{r run-prediction_scores}
# These are already in the saved "hamby44_eval.rda" file, so no need to re-run, but as an example of how to do it with a new method! 

#hamby44_eval <- prediction_scores(hamby44_eval, "rlo")
#hamby44_eval <- prediction_scores(hamby44_eval, "quadratic") 

hamby44_eval <- prediction_scores(hamby44_eval, "logistic4")
head(hamby44_eval)
```





```{r define-plot_scores}
plot_scores <- function(dataset, methods = NULL){
  colnames <- c()
  for (i in 1:length(methods)){
    colnames <- c(colnames, paste0("score_left_", methods[i]))
    colnames <- c(colnames, paste0("score_right_", methods[i]))
  }
  plot_df <- dataset %>% ungroup() %>% 
    select(colnames) %>%  
    gather(method, score) %>% 
    mutate(GrooveMethod = unlist(purrr::map(method, .f = function(x){
      strsplit(x, "_")[[1]][3]
        })), GrooveSide = unlist(purrr::map(method, .f = function(x){
      strsplit(x, "_")[[1]][2]
        }))
    )
  plot_df %>% 
    ggplot() + geom_boxplot(aes(x = GrooveSide, y = abs(score), fill = GrooveMethod)) + 
      theme_bw() + labs(x = "Groove Side", y = "Inaccuracy Score")
}

plot_scores_zoom <- function(dataset, zoom_cutoff, methods = NULL){
  colnames <- c()
  for (i in 1:length(methods)){
    colnames <- c(colnames, paste0("score_left_", methods[i]))
    colnames <- c(colnames, paste0("score_right_", methods[i]))
  }
  plot_df <- dataset %>% ungroup() %>% 
    select(colnames) %>%  
    gather(method, score) %>% 
    mutate(GrooveMethod = unlist(purrr::map(method, .f = function(x){
      strsplit(x, "_")[[1]][3]
        })), GrooveSide = unlist(purrr::map(method, .f = function(x){
      strsplit(x, "_")[[1]][2]
        }))
    ) %>% 
    filter(score < zoom_cutoff)
  plot_df %>% 
    ggplot() + geom_boxplot(aes(x = GrooveSide, y = abs(score), fill = GrooveMethod)) + 
      theme_bw() + labs(x = "Groove Side", y = "Inaccuracy Score")
  }
```

```{r plot_scores}
plot_scores(hamby44_eval, methods = c("rollapply", "rlo", "logistic", "logistic4"))

plot_scores_zoom(hamby44_eval, zoom_cutoff = 5000, methods = c("rollapply", "rlo", "logistic", "logistic4"))
```


```{r save-hamby44_eval}
saveRDS(hamby44_eval, "data/hamby44/hamby44_eval.rda")
```





## Metadata .csv for "problem lands"  

```{r prediction-metadata}
#hamby44_mdata <- hamby44_eval %>% ungroup() %>% select(1:4, dplyr::contains("score"))
#write.csv(hamby44_mdata, file = "data/hamby44/hamby44_mdata.csv")

hamby44_mdata <- read.csv(file = "data/hamby44/hamby44_mdata.csv")

```
