---
title: "Problem Description"
author: "Ulrike Genschel, Kiegan Rice"
output: html_document
---

#### Things to make sure we have loaded when working with this data:  

```{r, message = FALSE, warning = FALSE}
library(devtools)
library(tidyverse)
devtools::install_github("erichare/bulletr")
library(bulletr)
```

#### How can we identify the location of the grooves at the edge of a bullet land?  
_What information do we have?_  

1. 3D bullet scan data in x3p format: in a matrix, values are 'z' (height) values.  
2. 3D bullet scan data in data frame format: using `fortify_x3p` from Eric's `bulletr` package, we can get `x`, `y`, and `value` columns. These `x` and `y` have been converted to the actual micrometer location values instead of indices.  
3. Profiles of bullet lands   
    + For a specific value of `x`, we can plot the profile by plotting `z` vs the `y` values.  
  

```{r, warning = FALSE}
br411_df <- bulletr::fortify_x3p(br411)
br411_df %>% filter(between(x, 100, 110)) %>% ggplot() + geom_point(aes(x = y, y = value), size = .1)
```

4. `get_grooves` function in `bulletr` package. This is the current method of identifying the grooves at the edge of the bullet lands.  
    + 
    