---
title: "Groove Predictions: Impact on Bullet-Matching Algorithm"
output: html_document
---

# Hamby44  

## Rollapply Grooves  

```{r}
library(dplyr)
library(bulletxtrctr)
library(x3ptools)
library(randomForest)
library(ggplot2)
library(readr)
library(furrr)


bullets <- readRDS("../data/hamby44/hamby44_paper.rda")

## PUT A DUMMY "Y" = 0 into CCDATA
```

Extract Features   

```{r}

bullets <- bullets %>% mutate(
  sigs = purrr::map2(
    .x = ccdata, .y = grooves_rollapply, 
    .f = function(x, y) {
      cc_get_signature(
        ccdata = x, grooves = y, span1 = 0.75, span2 = 0.03)
    })
)
```


Align signatures and extract features   


```{r}
bullets$bulletland <- paste0(bullets$bullet,"-", bullets$land)
bullets$barrelbulletland <- paste0(bullets$barrel, "-", bullets$bullet, "-", bullets$land)
lands <- unique(bullets$barrelbulletland)
comparisons <- data.frame(
  expand.grid(land1 = lands, land2 = lands), stringsAsFactors = FALSE)

#cores <- function() { max(1, round(0.25*availableCores())) }
#plan(tweak(multiprocess, workers = cores))
comparisons <- comparisons %>% mutate(
  aligned = purrr::map2(.x = land1, .y = land2, .f = function(xx, yy) {
    land1 <- bullets$sigs[bullets$barrelbulletland == xx][[1]]
    land2 <- bullets$sigs[bullets$barrelbulletland == yy][[1]]
    land1$bullet <- "first-land"
    land2$bullet <- "second-land"
    
    sig_align(land1$sig, land2$sig)
  })
)


comparisons <- comparisons %>% mutate(
  ccf0 = aligned %>% 
    purrr::map_dbl(.f = function(x) extract_feature_ccf(x$lands)),
  lag0 = aligned %>% 
    purrr::map_dbl(.f = function(x) extract_feature_lag(x$lands)),
  D0 = aligned %>% 
    purrr::map_dbl(.f = function(x) extract_feature_D(x$lands)),
  length0 = aligned %>% 
    purrr::map_dbl(.f = function(x) extract_feature_length(x$lands)),
  overlap0 = aligned %>% 
    purrr::map_dbl(.f = function(x) extract_feature_overlap(x$lands))
)

comparisons <- comparisons %>% mutate(
  striae = aligned %>% purrr::map(.f = sig_cms_max, span = 75) 
)
comparisons <- comparisons %>% mutate(
  cms_per_mm = purrr::map2(striae, aligned, .f = function(s, a) {
    extract_feature_cms_per_mm(s$lines, a$lands, resolution=0.645)
  }),
  matches0 = striae %>% purrr::map_dbl(.f = function(s) {
    bulletxtrctr:::extract_helper_feature_n_striae(s$lines, type = "peak", match = TRUE)
  }),
  mismatches0 = striae %>% purrr::map_dbl(.f = function(s) {
    bulletxtrctr:::extract_helper_feature_n_striae(s$lines, type = "peak", match = FALSE)
  })
  
)

comparisons <- comparisons %>% mutate(
  bulletA = gsub("([1-2])-([1-6])","\\1",land1),
  bulletB = gsub("([1-2])-([1-6])","\\1",land2),
  landA = gsub("([1-2])-([1-6])","\\2",land1),
  landB = gsub("([1-2])-([1-6])","\\2",land2)
)

comparisons <- comparisons %>% mutate(
  features = purrr::map2(.x = aligned, .y = striae, .f = extract_features_all, resolution = 0.645)
)

comparisons <- comparisons %>% mutate(
  legacy_features = purrr::map(striae, extract_features_all_legacy, resolution = 0.645)
)

comparisons <- comparisons %>% tidyr::unnest(legacy_features) 
# scale features before using them in the random forest, legacy features can be used out of the box

comparisons$rfscore <- predict(bulletxtrctr::rtrees, newdata = comparisons, type = "prob")[,2]

head(comparisons)

saveRDS(comparisons, "../data/hamby44/comparisons_rollapply.rda")

```

Aligned signatures of two matching lands:

```{r}
subset(comparisons, land1=="2-4" & land2 =="1-2")$aligned[[1]]$lands %>% 
  tidyr::gather(sigs, value, sig1, sig2) %>% 
  ggplot(aes(x = x, y = value, colour = sigs)) + 
    geom_line() +
  theme_bw() +
  scale_color_brewer(palette = "Dark2")


```


Some features are based on aligned signatures:

```{r}
comparisons <- comparisons %>% mutate(
  ccf0 = aligned %>% 
    purrr::map_dbl(.f = function(x) extract_feature_ccf(x$lands)),
  lag0 = aligned %>% 
    purrr::map_dbl(.f = function(x) extract_feature_lag(x$lands)),
  D0 = aligned %>% 
    purrr::map_dbl(.f = function(x) extract_feature_D(x$lands)),
  length0 = aligned %>% 
    purrr::map_dbl(.f = function(x) extract_feature_length(x$lands)),
  overlap0 = aligned %>% 
    purrr::map_dbl(.f = function(x) extract_feature_overlap(x$lands))
)
head(comparisons)
```


Other features need an evaluation of striation marks between two aligned signatures:  

```{r}
comparisons <- comparisons %>% mutate(
  striae = aligned %>% purrr::map(.f = sig_cms_max, span = 75) 
)
comparisons <- comparisons %>% mutate(
  cms_per_mm = purrr::map2(striae, aligned, .f = function(s, a) {
    extract_feature_cms_per_mm(s$lines, a$lands, resolution=0.645)
  }),
  matches0 = striae %>% purrr::map_dbl(.f = function(s) {
    bulletxtrctr:::extract_helper_feature_n_striae(s$lines, type = "peak", match = TRUE)
  }),
  mismatches0 = striae %>% purrr::map_dbl(.f = function(s) {
    bulletxtrctr:::extract_helper_feature_n_striae(s$lines, type = "peak", match = FALSE)
  })
  
)

comparisons <- comparisons %>% mutate(
  bulletA = gsub("([1-2])-([1-6])","\\1",land1),
  bulletB = gsub("([1-2])-([1-6])","\\1",land2),
  landA = gsub("([1-2])-([1-6])","\\2",land1),
  landB = gsub("([1-2])-([1-6])","\\2",land2)
)
```


Extract Features


```{r}
comparisons <- comparisons %>% mutate(
  features = purrr::map2(.x = aligned, .y = striae, .f = extract_features_all, resolution = 0.645)
)

comparisons <- comparisons %>% mutate(
  legacy_features = purrr::map(striae, extract_features_all_legacy, resolution = 0.645)
)

comparisons <- comparisons %>% tidyr::unnest(legacy_features) 
# scale features before using them in the random forest, legacy features can be used out of the box


# quick visualization:
comparisons %>% 
  ggplot(aes(x = landA, y = landB, fill = ccf)) +
  geom_tile() +
  scale_fill_gradient2(low = "grey80", high = "darkorange", 
                       midpoint = 0.5) +
  facet_grid(bulletB~bulletA, labeller = "label_both") +
  xlab("Land A") +
  ylab("Land B") +
  theme(aspect.ratio = 1)

```


Get Score predictions for each land to land comparison

```{r}
comparisons$rfscore <- predict(bulletxtrctr::rtrees, newdata = comparisons, type = "prob")[,2]

head(comparisons)

saveRDS(comparisons, "../data/hamby44/comparisons_rollapply.rda")

comparisons_rollapply <- comparisons
rfscores_roll <- comparisons_rollapply %>% select(land1, land2, rfscore)
saveRDS(rfscores_roll, "../data/hamby44/rfscores_roll.rda")

comparisons_viz <- comparisons
comparisons_viz <- comparisons_viz %>% select(-c(landA, landB)) %>% separate(bulletA, c("barrelA", "bulletA", "landA"), "-") %>% separate(bulletB, c("barrelB", "bulletB", "landB"),"-")

head(comparisons_viz)


subset(comparisons_viz, land1 == "Barrel 1-Bullet 1-Land 1" & land2 == "Barrel 1-Bullet 2-Land 5")$aligned[[1]]$lands %>% 
  tidyr::gather(sigs, value, sig1, sig2) %>% 
  ggplot(aes(x = x, y = value, colour = sigs)) + 
    geom_line() +
  theme_bw() +
  scale_color_brewer(palette = "Dark2")

comparisons_viz %>% 
  filter(barrelA == "Barrel 10" & barrelB == "Barrel 10") %>%
  filter(bulletA == "Bullet 1" & bulletB == "Bullet 2") %>% 
  ggplot(aes(x = landA, y = landB, fill = rfscore)) +
  geom_tile() +
  scale_fill_gradient2(low = "grey80", high = "darkorange", 
                       midpoint = .5) +
  facet_grid(bulletB~bulletA, labeller = "label_both") +
  xlab("Land A") +
  ylab("Land B") +
  theme(aspect.ratio = 1) 
```


## Lassobasic Grooves  

```{r}
bullets_lb <- readRDS("../data/hamby44/hamby44_paper.rda")
bullets_lb <- bullets_lb %>% mutate(
  sigs = purrr::map2(
    .x = ccdata, .y = grooves_lassobasic, 
    .f = function(x, y) {
      cc_get_signature(
        ccdata = x, grooves = y, span1 = 0.75, span2 = 0.03)
    })
)
#head(bullets_lb)


bullets_lb$bulletland <- paste0(bullets_lb$bullet,"-", bullets_lb$land)
bullets_lb$barrelbulletland <- paste0(bullets_lb$barrel, "-", bullets_lb$bullet, "-", bullets_lb$land)
lands_lb <- unique(bullets_lb$barrelbulletland)
comparisons_lb <- data.frame(
  expand.grid(land1 = lands_lb, land2 = lands_lb), stringsAsFactors = FALSE)

comparisons_lb <- comparisons_lb %>% mutate(
  aligned = purrr::map2(.x = land1, .y = land2, .f = function(xx, yy) {
    land1 <- bullets_lb$sigs[bullets_lb$barrelbulletland == xx][[1]]
    land2 <- bullets_lb$sigs[bullets_lb$barrelbulletland == yy][[1]]
    land1$bullet <- "first-land"
    land2$bullet <- "second-land"
    
    sig_align(land1$sig, land2$sig)
  })
)

#head(comparisons_lb)

comparisons_lb <- comparisons_lb %>% mutate(
  ccf0 = aligned %>% 
    purrr::map_dbl(.f = function(x) extract_feature_ccf(x$lands)),
  lag0 = aligned %>% 
    purrr::map_dbl(.f = function(x) extract_feature_lag(x$lands)),
  D0 = aligned %>% 
    purrr::map_dbl(.f = function(x) extract_feature_D(x$lands)),
  length0 = aligned %>% 
    purrr::map_dbl(.f = function(x) extract_feature_length(x$lands)),
  overlap0 = aligned %>% 
    purrr::map_dbl(.f = function(x) extract_feature_overlap(x$lands))
)

comparisons_lb <- comparisons_lb %>% mutate(
  striae = aligned %>% purrr::map(.f = sig_cms_max, span = 75) 
)
comparisons_lb <- comparisons_lb %>% mutate(
  cms_per_mm = purrr::map2(striae, aligned, .f = function(s, a) {
    extract_feature_cms_per_mm(s$lines, a$lands, resolution=0.645)
  }),
  matches0 = striae %>% purrr::map_dbl(.f = function(s) {
    bulletxtrctr:::extract_helper_feature_n_striae(s$lines, type = "peak", match = TRUE)
  }),
  mismatches0 = striae %>% purrr::map_dbl(.f = function(s) {
    bulletxtrctr:::extract_helper_feature_n_striae(s$lines, type = "peak", match = FALSE)
  })
  
)

comparisons_lb <- comparisons_lb %>% mutate(
  bulletA = gsub("([1-2])-([1-6])","\\1",land1),
  bulletB = gsub("([1-2])-([1-6])","\\1",land2),
  landA = gsub("([1-2])-([1-6])","\\2",land1),
  landB = gsub("([1-2])-([1-6])","\\2",land2)
)

comparisons_lb <- comparisons_lb %>% mutate(
  features = purrr::map2(.x = aligned, .y = striae, .f = extract_features_all, resolution = 0.645)
)

comparisons_lb <- comparisons_lb %>% mutate(
  legacy_features = purrr::map(striae, extract_features_all_legacy, resolution = 0.645)
)

comparisons_lb <- comparisons_lb %>% tidyr::unnest(legacy_features) 
# scale features before using them in the random forest, legacy features can be used out of the box
comparisons_lb$rfscore <- predict(bulletxtrctr::rtrees, newdata = comparisons_lb, type = "prob")[,2]

head(comparisons_lb)

saveRDS(comparisons_lb, "../data/hamby44/comparisons_lassobasic.rda")

rfscores_lb <- comparisons_lb %>% select(land1, land2, rfscore)
saveRDS(rfscores_lb, "../data/hamby44/rfscores_lb.rda")
```






## Lassofull Grooves  

```{r}
bullets_lf <- readRDS("../data/hamby44/hamby44_paper.rda")
bullets_lf <- bullets_lf %>% mutate(
  sigs = purrr::map2(
    .x = ccdata, .y = grooves_lassofull, 
    .f = function(x, y) {
      cc_get_signature(
        ccdata = x, grooves = y, span1 = 0.75, span2 = 0.03)
    })
)
#head(bullets_lf)


bullets_lf$bulletland <- paste0(bullets_lf$bullet,"-", bullets_lf$land)
bullets_lf$barrelbulletland <- paste0(bullets_lf$barrel, "-", bullets_lf$bullet, "-", bullets_lf$land)
lands_lf <- unique(bullets_lf$barrelbulletland)
comparisons_lf <- data.frame(
  expand.grid(land1 = lands_lf, land2 = lands_lf), stringsAsFactors = FALSE)

comparisons_lf <- comparisons_lf %>% mutate(
  aligned = purrr::map2(.x = land1, .y = land2, .f = function(xx, yy) {
    land1 <- bullets_lf$sigs[bullets_lf$barrelbulletland == xx][[1]]
    land2 <- bullets_lf$sigs[bullets_lf$barrelbulletland == yy][[1]]
    land1$bullet <- "first-land"
    land2$bullet <- "second-land"
    
    sig_align(land1$sig, land2$sig)
  })
)

#head(comparisons_lf)

comparisons_lf <- comparisons_lf %>% mutate(
  ccf0 = aligned %>% 
    purrr::map_dbl(.f = function(x) extract_feature_ccf(x$lands)),
  lag0 = aligned %>% 
    purrr::map_dbl(.f = function(x) extract_feature_lag(x$lands)),
  D0 = aligned %>% 
    purrr::map_dbl(.f = function(x) extract_feature_D(x$lands)),
  length0 = aligned %>% 
    purrr::map_dbl(.f = function(x) extract_feature_length(x$lands)),
  overlap0 = aligned %>% 
    purrr::map_dbl(.f = function(x) extract_feature_overlap(x$lands))
)

comparisons_lf <- comparisons_lf %>% mutate(
  striae = aligned %>% purrr::map(.f = sig_cms_max, span = 75) 
)
comparisons_lf <- comparisons_lf %>% mutate(
  cms_per_mm = purrr::map2(striae, aligned, .f = function(s, a) {
    extract_feature_cms_per_mm(s$lines, a$lands, resolution=0.645)
  }),
  matches0 = striae %>% purrr::map_dbl(.f = function(s) {
    bulletxtrctr:::extract_helper_feature_n_striae(s$lines, type = "peak", match = TRUE)
  }),
  mismatches0 = striae %>% purrr::map_dbl(.f = function(s) {
    bulletxtrctr:::extract_helper_feature_n_striae(s$lines, type = "peak", match = FALSE)
  })
  
)

comparisons_lf <- comparisons_lf %>% mutate(
  bulletA = gsub("([1-2])-([1-6])","\\1",land1),
  bulletB = gsub("([1-2])-([1-6])","\\1",land2),
  landA = gsub("([1-2])-([1-6])","\\2",land1),
  landB = gsub("([1-2])-([1-6])","\\2",land2)
)

comparisons_lf <- comparisons_lf %>% mutate(
  features = purrr::map2(.x = aligned, .y = striae, .f = extract_features_all, resolution = 0.645)
)

comparisons_lf <- comparisons_lf %>% mutate(
  legacy_features = purrr::map(striae, extract_features_all_legacy, resolution = 0.645)
)

comparisons_lf <- comparisons_lf %>% tidyr::unnest(legacy_features) 
# scale features before using them in the random forest, legacy features can be used out of the box
comparisons_lf$rfscore <- predict(bulletxtrctr::rtrees, newdata = comparisons_lf, type = "prob")[,2]

head(comparisons_lf)

saveRDS(comparisons_lf, "../data/hamby44/comparisons_lassofull.rda")
rfscores_lf <- comparisons_lf %>% select(land1, land2, rfscore)
saveRDS(rfscores_lf, "../data/hamby44/rfscores_lf.rda")

```








## BCP Grooves  

```{r}
bullets_bcp <- readRDS("../data/hamby44/hamby44_paper.rda")
bullets_bcp <- bullets_bcp %>% mutate(
  sigs = purrr::map2(
    .x = ccdata, .y = grooves_bcp, 
    .f = function(x, y) {
      cc_get_signature(
        ccdata = x, grooves = y, span1 = 0.75, span2 = 0.03)
    })
)
#head(bullets_bcp)


bullets_bcp$bulletland <- paste0(bullets_bcp$bullet,"-", bullets_bcp$land)
bullets_bcp$barrelbulletland <- paste0(bullets_bcp$barrel, "-", bullets_bcp$bullet, "-", bullets_bcp$land)
lands_bcp <- unique(bullets_bcp$barrelbulletland)
comparisons_bcp <- data.frame(
  expand.grid(land1 = lands_bcp, land2 = lands_bcp), stringsAsFactors = FALSE)

######
comparisons_bcp <- comparisons_bcp %>% mutate(
  aligned = purrr::map2(.x = land1, .y = land2, .f = function(xx, yy) {
    land1 <- bullets_bcp$sigs[bullets_bcp$barrelbulletland == xx][[1]]
    land2 <- bullets_bcp$sigs[bullets_bcp$barrelbulletland == yy][[1]]
    land1$bullet <- "first-land"
    land2$bullet <- "second-land"
    
    sig_align(land1$sig, land2$sig)
  })
)

#head(comparisons_lf)

comparisons_bcp <- comparisons_bcp %>% mutate(
  ccf0 = aligned %>% 
    purrr::map_dbl(.f = function(x) extract_feature_ccf(x$lands)),
  lag0 = aligned %>% 
    purrr::map_dbl(.f = function(x) extract_feature_lag(x$lands)),
  D0 = aligned %>% 
    purrr::map_dbl(.f = function(x) extract_feature_D(x$lands)),
  length0 = aligned %>% 
    purrr::map_dbl(.f = function(x) extract_feature_length(x$lands)),
  overlap0 = aligned %>% 
    purrr::map_dbl(.f = function(x) extract_feature_overlap(x$lands))
)

comparisons_bcp <- comparisons_bcp %>% mutate(
  striae = aligned %>% purrr::map(.f = sig_cms_max, span = 75) 
)
comparisons_bcp <- comparisons_bcp %>% mutate(
  cms_per_mm = purrr::map2(striae, aligned, .f = function(s, a) {
    extract_feature_cms_per_mm(s$lines, a$lands, resolution=0.645)
  }),
  matches0 = striae %>% purrr::map_dbl(.f = function(s) {
    bulletxtrctr:::extract_helper_feature_n_striae(s$lines, type = "peak", match = TRUE)
  }),
  mismatches0 = striae %>% purrr::map_dbl(.f = function(s) {
    bulletxtrctr:::extract_helper_feature_n_striae(s$lines, type = "peak", match = FALSE)
  })
  
)

comparisons_bcp <- comparisons_bcp %>% mutate(
  bulletA = gsub("([1-2])-([1-6])","\\1",land1),
  bulletB = gsub("([1-2])-([1-6])","\\1",land2),
  landA = gsub("([1-2])-([1-6])","\\2",land1),
  landB = gsub("([1-2])-([1-6])","\\2",land2)
)

comparisons_bcp <- comparisons_bcp %>% mutate(
  features = purrr::map2(.x = aligned, .y = striae, .f = extract_features_all, resolution = 0.645)
)

comparisons_bcp <- comparisons_bcp %>% mutate(
  legacy_features = purrr::map(striae, extract_features_all_legacy, resolution = 0.645)
)

comparisons_bcp <- comparisons_bcp %>% tidyr::unnest(legacy_features) 
# scale features before using them in the random forest, legacy features can be used out of the box
comparisons_bcp$rfscore <- predict(bulletxtrctr::rtrees, newdata = comparisons_bcp, type = "prob")[,2]

head(comparisons_bcp)

saveRDS(comparisons_bcp, "../data/hamby44/comparisons_bcp.rda")
rfscores_bcp <- comparisons_bcp %>% select(land1, land2, rfscore)
saveRDS(rfscores_bcp, "../data/hamby44/rfscores_bcp.rda")

```






## Manual Grooves  


```{r}
bullets_manual <- readRDS("../data/hamby44/hamby44_paper.rda")

bullets_manual <- bullets_manual %>% mutate(
  sigs = purrr::map2(
    .x = ccdata, .y = grooves_kr1, 
    .f = function(x, y) {
      cc_get_signature(
        ccdata = x, grooves = y, span1 = 0.75, span2 = 0.03)
    })
)
#head(bullets_manual)


bullets_manual$bulletland <- paste0(bullets_manual$bullet,"-", bullets_manual$land)
bullets_manual$barrelbulletland <- paste0(bullets_manual$barrel, "-", bullets_manual$bullet, "-", bullets_manual$land)
lands_manual <- unique(bullets_manual$barrelbulletland)
comparisons_manual <- data.frame(
  expand.grid(land1 = lands_manual, land2 = lands_manual), stringsAsFactors = FALSE)

######
comparisons_manual <- comparisons_manual %>% mutate(
  aligned = purrr::map2(.x = land1, .y = land2, .f = function(xx, yy) {
    land1 <- bullets_manual$sigs[bullets_manual$barrelbulletland == xx][[1]]
    land2 <- bullets_manual$sigs[bullets_manual$barrelbulletland == yy][[1]]
    land1$bullet <- "first-land"
    land2$bullet <- "second-land"
    
    sig_align(land1$sig, land2$sig)
  })
)

#head(comparisons_manual)

comparisons_manual <- comparisons_manual %>% mutate(
  ccf0 = aligned %>% 
    purrr::map_dbl(.f = function(x) extract_feature_ccf(x$lands)),
  lag0 = aligned %>% 
    purrr::map_dbl(.f = function(x) extract_feature_lag(x$lands)),
  D0 = aligned %>% 
    purrr::map_dbl(.f = function(x) extract_feature_D(x$lands)),
  length0 = aligned %>% 
    purrr::map_dbl(.f = function(x) extract_feature_length(x$lands)),
  overlap0 = aligned %>% 
    purrr::map_dbl(.f = function(x) extract_feature_overlap(x$lands))
)

comparisons_manual <- comparisons_manual %>% mutate(
  striae = aligned %>% purrr::map(.f = sig_cms_max, span = 75) 
)
comparisons_manual <- comparisons_manual %>% mutate(
  cms_per_mm = purrr::map2(striae, aligned, .f = function(s, a) {
    extract_feature_cms_per_mm(s$lines, a$lands, resolution=0.645)
  }),
  matches0 = striae %>% purrr::map_dbl(.f = function(s) {
    bulletxtrctr:::extract_helper_feature_n_striae(s$lines, type = "peak", match = TRUE)
  }),
  mismatches0 = striae %>% purrr::map_dbl(.f = function(s) {
    bulletxtrctr:::extract_helper_feature_n_striae(s$lines, type = "peak", match = FALSE)
  })
  
)

comparisons_manual <- comparisons_manual %>% mutate(
  bulletA = gsub("([1-2])-([1-6])","\\1",land1),
  bulletB = gsub("([1-2])-([1-6])","\\1",land2),
  landA = gsub("([1-2])-([1-6])","\\2",land1),
  landB = gsub("([1-2])-([1-6])","\\2",land2)
)

comparisons_manual <- comparisons_manual %>% mutate(
  features = purrr::map2(.x = aligned, .y = striae, .f = extract_features_all, resolution = 0.645)
)

comparisons_manual <- comparisons_manual %>% mutate(
  legacy_features = purrr::map(striae, extract_features_all_legacy, resolution = 0.645)
)

comparisons_manual <- comparisons_manual %>% tidyr::unnest(legacy_features) 
# scale features before using them in the random forest, legacy features can be used out of the box
comparisons_manual$rfscore <- predict(bulletxtrctr::rtrees, newdata = comparisons_manual, type = "prob")[,2]

head(comparisons_manual)

saveRDS(comparisons_manual, "../data/hamby44/comparisons_manual.rda")
rfscores_manual <- comparisons_manual %>% select(land1, land2, rfscore)
saveRDS(rfscores_manual, "../data/hamby44/rfscores_manual.rda")

```

